---
title: "SNA_pred"
author: "Manu"
date: "April 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(igraph)

load("C:/Users/ms52/Desktop/Adv Social Networks/DataSets/Glasgow_data/Glasgow-demographic.RData")
load("C:/Users/ms52/Desktop/Adv Social Networks/DataSets/Glasgow_data/Glasgow-friendship.RData")
# try and viz the friendship networks - 

friendship.1 = ifelse( friendship.1 == 10, 0, friendship.1)
#friendship.1 = ifelse( friendship.1 == 2, 1, friendship.1)
g1 <- graph_from_adjacency_matrix(friendship.1)


# friendship.2 = ifelse( friendship.2 == 10, 0, friendship.2)
# friendship.2 = ifelse( friendship.2 == 2, 1, friendship.2)
# g2 <- graph_from_adjacency_matrix(friendship.2)


friendship.3 = ifelse( friendship.3 == 10, 0, friendship.3)
#friendship.3 = ifelse( friendship.3 == 2, 1, friendship.3)
g3 <- graph_from_adjacency_matrix(friendship.3)

```
Attach all the various attributes to the friendship graph-

```{r}

plot(g1, edge.arrow.size=.5, vertex.color="gold", vertex.size=15, 

     vertex.frame.color="gray", vertex.label.color="black", 

     vertex.label.cex=0.5, vertex.label.dist=0, edge.curved=0.2) 
layout <- layout_nicely(g1, dim = 3)

sex_color = get.vertex.attribute(g1,"sex")
sex_color[sex_color == 2] = "red"
sex_color[sex_color == 1] = "blue"

plot(g1, 
     layout=layout, 
     vertex.color=sex_color, 
     vertex.label=NA, 
     edge.arrow.size=.5,
     main='School Network by sex of student')


 g1 <- graph_from_adjacency_matrix(friendship.1)
 summary(g1)

tobacco = as.data.frame(tobacco)
tobacco$id = rownames(tobacco)

fam = as.data.frame(familysmoking)
fam$id = rownames(fam)

age = as.data.frame(age)
age$id = rownames(age)
sex.F = as.data.frame(sex.F)
sex.F$id = rownames(sex.F)


V(g1)$age=as.numeric(age$age[match(V(g1)$name,age$id)])
V(g1)$sex=as.numeric(sex.F$sex.F[match(V(g1)$name,sex.F$id)])
V(g1)$smoke=as.numeric(tobacco$t1[match(V(g1)$name,tobacco$id)])
V(g1)$fam.smoke=as.numeric(fam$smoking.at.home[match(V(g1)$name,fam$id)])
# 
# 
#
```

created graph G1 and now similarly we will create graph G2 .

This is a prelim model and I am only training on a handful of attributes present in the mater dataset. Namely - age, sex and family history of smoking. (alongside network information of course.)


```{r}



 g3 <- graph_from_adjacency_matrix(friendship.3)
 summary(g3)

V(g3)$age=as.numeric(age$age[match(V(g3)$name,age$id)])
V(g3)$sex=as.numeric(sex.F$sex.F[match(V(g3)$name,sex.F$id)])
V(g3)$smoke=as.numeric(tobacco$t1[match(V(g3)$name,tobacco$id)])
V(g3)$fam.smoke=as.numeric(fam$smoking.at.home[match(V(g3)$name,fam$id)])


```


## both networks are ready- now to run a simple linear regression model.

```{r}

total = merge(age, sex.F, by = "id")
total = merge(total, fam, by = "id")
total = total[,-c(5,6)]
total = merge(total, tobacco, by = "id")
total = total[,-c(6,7)]

total$smoking.at.home = ifelse(total$smoking.at.home ==2, 1, 0)
total$sex.F = ifelse(total$sex.F ==1, 1, 0)


head(total)

str(total)

total$sex.F = as.factor(total$sex.F)
total$smoking.at.home = as.factor(total$smoking.at.home)
```
With the clean dataset now to make a linear regression model :

Objective here is to compare the performance of a linear model with that of a model which has additional network information. Because none of the factors are invariant in time in this smaller subset - (i only have sex, age and family smoking- which changes insignificantly over the span of the study). If I trained my model at one point of time and used it to predict in the future- I will get almost perfect predictions (the training and testing datasets are almost the same!!)

So to avoid that issue I will partition the dataset into training and testing at time t1.
```{r}

 lm1 = lm(t1~age+sex.F+smoking.at.home, data = total)

summary(lm1)

library(car)

pairs(t1~age+sex.F+smoking.at.home, data = total,
   main="Relationship between various elements")

avPlots(lm1)

pred = total[,1:4]
smoking = as.data.frame(predict(lm1,pred))

head(smoking)
library(caret)     
library(mlbench)

rmse <- function(error)
{
    sqrt(mean(error^2))
}

rmse(lm1$residuals)
postResample(pred = smoking, obs = tobacco$t1 )
#confusionMatrix(data = smoking$`predict(lm1, pred)`, reference = tobacco$t1, mode = "prec_recall")

```

Trying to model the same with network data
```{r}

n = as.data.frame(unlist(ego(g1,1, 1, mode = "all")))
n$neigh = rownames(n)
n= n[,-1]
n
total$weight = ifelse(total$id %in% n , 3,1)



```
