---
title: "Prediction Mini Project"
author: "Manu"
date: "April 19, 2017"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
## Objective

Just as a reminder in my previous proposal I raised the following questions:

Hypothesis 1a: The study of network structure in conjunction with survey information
can better predict or identify the individuals who are at the maximum risk of behaviour adoption
as compared to predictions made using surveys alone. (i.e. social networks in such a context
add significant value)

Hypothesis 1b: In the context of predictive analysis, model complexity will add value
i.e. machine learning algorithms such as KNN should perform better than simple regression
models.

Hypothesis 2: Not just the closest friends (i.e. degree one friendships) but extended friend
circles (i.e. degree two or three or more) also impacts an individual’s decision to engage in
undesirable behaviour


Now as part of the miniproject I want to test the validity of hypothesis 1a. ( I also have some interesting results for hypothesis 2 but it needs more cleaning up). 

####Loading the cleaned and preprocessed datasets - 

```{r}
library(igraph)

load("C:/Users/SONY/Desktop/Adv Soc Networks/Data/Glasgow-friendship.RData")
load("C:/Users/SONY/Desktop/Adv Soc Networks/Data/Glasgow-demographic.RData")
load("C:/Users/SONY/Desktop/Adv Soc Networks/Data/Glasgow-substances.RData")
load("C:/Users/SONY/Desktop/Adv Soc Networks/Data/Glasgow-various.RData")
# try and viz the friendship networks - 

friendship.1 = ifelse( friendship.1 == 10, 0, friendship.1)
g1 <- graph_from_adjacency_matrix(friendship.1)
friendship.3 = ifelse( friendship.3 == 10, 0, friendship.3)
g3 <- graph_from_adjacency_matrix(friendship.3)

```

Attaching some node attributes to the graph  
```{r}

g1 <- graph_from_adjacency_matrix(friendship.1)
#summary(g1)

tobacco = as.data.frame(tobacco)
tobacco$id = rownames(tobacco)

fam = as.data.frame(familysmoking)
fam$id = rownames(fam)

age = as.data.frame(age)
age$id = rownames(age)
sex.F = as.data.frame(sex.F)
sex.F$id = rownames(sex.F)

V(g1)$age=as.numeric(age$age[match(V(g1)$name,age$id)])
V(g1)$sex=as.numeric(sex.F$sex.F[match(V(g1)$name,sex.F$id)])
V(g1)$smoke=as.numeric(tobacco$t1[match(V(g1)$name,tobacco$id)])
V(g1)$fam.smoke=as.numeric(fam$smoking.at.home[match(V(g1)$name,fam$id)])
```

As this is a preliminary model I am taking only a handful of attributes to start off with. The actual dataset has many other factors which I will incorporate in the final project. The mini project is also reflective of the way I thought about the problem so I think it should be easy to follow and intuitive.


##Simple model without network inputs

When I think about smoking in adolescents some factors that seem most predictive are age, sex and whether a family member (parents or sibling) smokes. Thus to start off the analysis I considered only these three factors to understand how well they capture the smoking behaviour.  
```{r}

total = merge(age, sex.F, by = "id")
total = merge(total, fam, by = "id")
total = total[,-c(5,6)]
#total = merge(total, tobacco, by = "id")
total = total[,-c(6,7)]

total$smoking.at.home = ifelse(total$smoking.at.home ==2, 1, 0)
total$sex.F = ifelse(total$sex.F ==1, 1, 0)
total$sex.F = as.factor(total$sex.F)
#total$smoking.at.home = as.factor(total$smoking.at.home)
# adding other factors into regression

#average distance from friends


#in a romantic relationship - unformtunately who they were dating is not given  

romantic1 = romantic
colnames(romantic) = c ("romantic1", "romantic2", "romantic3")
romantic[romantic ==1] = 0
romantic[romantic ==2] =1
romantic = ifelse(is.na(romantic), 0,romantic)
summary(romantic)
romantic = as.data.frame(romantic); romantic$id = row.names(romantic)
#money available to students
colnames(money) = c ("money1", "money2", "money3")
money[money < 0] <- 0
money = as.data.frame(money); money$id = row.names(money)
#substances used

colnames(tobacco) = c("smoking1", "smoking2", "smoking3")
tobacco = as.data.frame(tobacco); tobacco$id = row.names(tobacco)

colnames(alcohol) = c("alcohol1", "alcohol2", "alcohol3")
alcohol = as.data.frame(alcohol); alcohol$id = row.names(alcohol)

colnames(cannabis) = c("cannabis1", "cannabis2", "cannabis3")
cannabis = as.data.frame(cannabis); cannabis$id = row.names(cannabis)

colnames(dist.school) = c("dist.school1", "dist.school2", "dist.school3")
dist.school = as.data.frame(dist.school); dist.school$id = row.names(dist.school)


# average distance from degree 1 friends




head(total)
#total = total[,-5]
total1 = total

total = merge(total, romantic, by ="id")
total = merge(total, money, by ="id")
total = merge(total, dist.school, by ="id")
total = merge(total, alcohol, by ="id")
total = merge(total, cannabis, by ="id")
total = merge(total, tobacco, by ="id")

head(total)
```

```{r}

lm1 = lm(smoking1~age+sex.F+smoking.at.home+romantic1+money1+dist.school1, data = total)
summary(lm1)



lm2 = lm(smoking1~age+sex.F+smoking.at.home+romantic1+money1+dist.school1+alcohol1+cannabis1, data = total)

summary(lm2)

library(car)

avPlots(lm2)
pred = total[,1:4]
smoking = as.data.frame(predict(lm1,pred))
library(caret)     
library(mlbench)

rmse <- function(error)
{
  sqrt(mean(error^2))
}

rmse(lm1$residuals)
postResample(pred = smoking, obs = tobacco$t1 )
#plot(lm1)

#confusionMatrix(data = smoking$`predict(lm1, pred)`, reference = tobacco$t1, mode = "prec_recall")

```
The linear regression showed me some very surprising results. While age and gender seem to affect smoking. The presence of a family member who smokes does not  ! (Will need to dig deeper into the issue). But it seems to me looking at the residuals and R sq values that this model is not doing well at all. 

###Model with Network inputs

Now I want to throw in some network information in the mix to see if this improves the model performance. I want to test if a network metric such as centrality will affect smoking behaviour. Do more central and popular students take up smoking. To test this out I calculate 4 different types if centralities - Closeness centrality, Betweeness centrality, Bonancich and eigenvector centrality. 

In the model below I run the regression with the addition of these network measures. 

```{r}

bc = as.data.frame(betweenness(g1)); bc$id = rownames(bc)
bon=  as.data.frame (bonpow(g1));bon$id = rownames(bon)
close = as.data.frame(closeness(g1)); close$id = rownames(close)
eigen = as.data.frame(eigen_centrality(g1)$vector)
eigen$id = rownames(eigen)
colnames(eigen)[1]="eigen_cent(g1)"

total = merge(total, bon, by ="id")
total = merge(total, bc, by ="id")
total = merge(total, close, by ="id")
total = merge(total, eigen, by ="id")
head(total)
total$id = rownames(total);total=total[,-1]
lm2 = lm(t1~ age+sex.F+smoking.at.home+`eigen_cent(g1)`+`betweenness(g1)`+`closeness(g1)`+`bonpow(g1)`, data = total)
summary(lm2)

```
Unfortunately none of the centrality measures have any impact on the smoking measures. Thus thinking about the problem as I would in real life, I feel an adolescents smoking behaviour is going to be predicted very strongly by their friends. I thus create a new network metric called "smoking index". This is defined loosely as the summation of number of smokers in close friends network (degree 1) divided by total number of friends 
Trying to model the same with network data
```{r}
nodes = list(total$id)
i = 0
s = 0
for (i in 1:160){
n = as.data.frame(unlist(ego(g1,1, i, mode = "all")))
n$neigh = rownames(n)
n= n[-1,-1]
n
s[i] = sum(vertex_attr(g1, "smoke", index = n))/degree(g1,i)
s
}

total$smoking_degree1 = s

lm3 = lm(smoking1~age+sex.F+smoking.at.home+romantic1+money1+dist.school1+smoking_degree1, data = total)

summary(lm3)
avPlots(lm3)
```


##calculating the indices for degree 2

```{r}

nodes = list(total$id)
i = 0
s = 0
for (i in 1:160){
n1 = as.data.frame(unlist(ego(g1,1, i, mode = "all")))
n1$neigh = rownames(n1)
n1= n1[-1,-1]
n1
n2 = as.data.frame(unlist(ego(g1,2, i, mode = "all")))
n2$neigh = rownames(n2)
n2= n2[-1,-1]
n2

n2 = n2[!n2 %in% n1]

s[i] = sum(vertex_attr(g1, "smoke", index = n2), na.rm = TRUE)/length(n2)
s
}

total$smoking_degree2 = s

lm4 = lm(smoking1~age+sex.F+smoking.at.home+romantic1+money1+dist.school1+smoking_degree2, data = total)

summary(lm4)

lm5 = lm(smoking1~age+sex.F+smoking.at.home+romantic1+money1+dist.school1+smoking_degree1+smoking_degree2, data = total)

summary(lm5)



```


```{r}

nodes = list(total$id)
i = 0
s = 0
for (i in 1:160){
n1 = as.data.frame(unlist(ego(g1,1, i, mode = "all")))
n1$neigh = rownames(n1)
n1= n1[-1,-1]
n1
n2 = as.data.frame(unlist(ego(g1,2, i, mode = "all")))
n2$neigh = rownames(n2)
n2= n2[-1,-1]
n2

n3 = as.data.frame(unlist(ego(g1,3, i, mode = "all")))
n3$neigh = rownames(n3)
n3= n3[-1,-1]
n3

n2 = n2[!n2 %in% n1]
n3 = n3[!n3 %in% n2]
s[i] = sum(vertex_attr(g1, "smoke", index = n3), na.rm = TRUE)/length(n3)
s
}

total$smoking_degree3 = s



lm6 = lm(smoking1~age+sex.F+smoking.at.home+romantic1+money1+dist.school1+smoking_degree1+smoking_degree2+smoking_degree3, data = total)

summary(lm6)

```
Fortunately model 3 fits the data much better than before. The residuals are much smaller and R sq has improved. Although this is not the complete story it goes on to show that I am progressing in the correct direction and including network data to a prediction problem certainly adds significant value.


model with all the predictors
```{r}
lm7 = lm(smoking1~age+sex.F+smoking.at.home+romantic1+money1+dist.school1+alcohol1+cannabis1+smoking_degree1+smoking_degree2+smoking_degree3, data = total)

summary(lm7)

```
#corrplot

```{r}
library(corrplot)
#total = total[,-23]
total$smoking.at.home = as.numeric(total$smoking.at.home)
temp = total[ , names(total)%in%c("age","smoking.at.home","romantic1","money1","dist.school1","alcohol1","cannabis1","smoking_degree1","smoking_degree2","smoking_degree3", "smoking1")]

m = cor(temp, use="complete.obs", method="pearson") 
corrplot(m, method="circle")
```

confusion matrix and sensitivity analysis

```{r}






````





###Next Steps

This was just my first attempt at understanding smoking using this dataset. In the subsequent iterations I will specify my model more explicitly. Instead of using simple regression I think the longitudinal logistic regression where smoking at any given time period (t+1) is a function of various demographic/network attributes such as age, sex, alcohol/marijuana consumption, , friends who smoke, distance between friends, relationship status, money availability, leisure activity preferences and more

I hope the use of a time lagged dependant variable eliminates serial correlation in the errors. The next steps which I am most keen on working are- if any other interesting network measures can be constructed from the data given. I also plan to generate smoking indices for degree 2,3... friends and weighing these columns differently to improve model performance. 
 
 Thus for now I would like ideas on which models you would like to see competing against each other and interesting network measures that can be constructed. I would also like to hear your thoughts on choosing weights for different degrees of seperation. At present my best idea is to run the model with different weight combinations a couple time and choose the weights which perform best, but there is going to be no theoretical backing for these. For the models I plan to use two parametric and one non parametric model and compare their performances.
 The final outcome will be a confusion matrix which will have results from the best performing predictive model and will give way to a little qualitative exploration (i.e. why do some cases predicted to be at highest risk of smoking do not take up smoking vs. some cases predicted to not have any inclination to smoke, take it up). Although the data that I have may limit how much I can understand qualitatively.  
