---
title: "Prelim_SNA"
author: "Manu"
date: "February 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading and vizulaizing the data sources:

The objective of the study is to move beyond linear, logistic regression and leverage the network structure to make predictions about peoples behaviour and choices. At baseline we can compare predictions from various parametric models with my proposed non- parametric modelling methodology. 


Simplifications made connections that did not exist (because the students moved or came in from different schools have been counted as null relationships i.e 0). The graph will be directed and weighted. 

The values are recoded as :

0- No friendship
1- Just a friend 
2- Best friend

For now we will use a non weighted directed graph . Just the presence of a friendship will be denoted as 1.

G1 G2 and G3 are the three friendship graphs at three points in time. 
```{r}
library(igraph)

load("C:/Users/ms52/Desktop/Adv Social Networks/DataSets/Glasgow_data/Glasgow-demographic.RData")
load("C:/Users/ms52/Desktop/Adv Social Networks/DataSets/Glasgow_data/Glasgow-friendship.RData")
# try and viz the friendship networks - 

friendship.1 = ifelse( friendship.1 == 10, 0, friendship.1)
friendship.1 = ifelse( friendship.1 == 2, 1, friendship.1)
g1 <- graph_from_adjacency_matrix(friendship.1)


friendship.2 = ifelse( friendship.2 == 10, 0, friendship.2)
friendship.2 = ifelse( friendship.2 == 2, 1, friendship.2)
g2 <- graph_from_adjacency_matrix(friendship.2)


friendship.3 = ifelse( friendship.3 == 10, 0, friendship.3)
friendship.3 = ifelse( friendship.3 == 2, 1, friendship.3)
g3 <- graph_from_adjacency_matrix(friendship.3)

```
Attach all the various attributes to the friendship graph-

```{r}

plot(g1, edge.arrow.size=.5, vertex.color="gold", vertex.size=15, 

     vertex.frame.color="gray", vertex.label.color="black", 

     vertex.label.cex=0.5, vertex.label.dist=0, edge.curved=0.2) 
layout <- layout_nicely(g1, dim = 3)

sex_color = get.vertex.attribute(g1,"sex")
sex_color[sex_color == 2] = "red"
sex_color[sex_color == 1] = "blue"

plot(g1, 
     layout=layout, 
     vertex.color=sex_color, 
     vertex.label=NA, 
     edge.arrow.size=.5,
     main='School Network by sex of student')


 g1 <- graph_from_adjacency_matrix(friendship.1)
 summary(g1)
# alcohol = as.data.frame(alcohol)
# alcohol$id = rownames(alcohol)
# tobacco = as.data.frame(tobacco)
# tobacco$id = rownames(tobacco)
# cannabis = as.data.frame(cannabis)
# cannabis$id = rownames(cannabis)
age = as.data.frame(age)
age$id = rownames(age)
sex.F = as.data.frame(sex.F)
sex.F$id = rownames(sex.F)
# dist.school = as.data.frame(dist.school)
# dist.school$id = rownames(dist.school)
# distance.1 = as.data.frame(distance.1)
# distance.1$id = rownames(distance.1)
# leisure1 = as.data.frame(leisure1)
# leisure1$id = rownames(leisure1)
# music1 = as.data.frame(music1)
# music1$id = rownames(music1)
# familysmoking = as.data.frame(familysmoking)
# familysmoking$id = rownames(familysmoking)
# romantic = as.data.frame(romantic)
# romantic$id = rownames(romantic)
# money = as.data.frame(money)
# money$id = rownames(money)
# 
# other = Reduce(function(x, y) merge(x, y, all=TRUE), list(leisure1,music1 , familysmoking))
# 
# 
# # alcohol.1 = alcohol[,c("t1", "id")]
# # colnames(alcohol.1) = c("alcohol", "id")
# # 
# # vertex_attr(g1, index=alcohol$id) <- alcohol.1
# # V(net)$Sex=as.character(a$Sex[match(V(net)$name,a$Bird.ID)])
# V(g1)$alcohol=as.numeric(alcohol$t1[match(V(g1)$name,alcohol$id)])
# V(g1)$tobacco=as.numeric(tobacco$t1[match(V(g1)$name,tobacco$id)])
# V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])

V(g1)$age=as.numeric(age$age[match(V(g1)$name,age$id)])
V(g1)$sex=as.numeric(sex.F$sex.F[match(V(g1)$name,sex.F$id)])

# V(g1)$dist.school=as.numeric(dist.school$t1[match(V(g1)$name,dist.school$id)])
# V(g1)$distance.1=as.numeric(distance.1$t1[match(V(g1)$name,distance.1$id)])
# V(g1)$smoking.at.home=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
# V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
# V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
# V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
# V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
# V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
# V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
# V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
# 
# 
# g1 = set_vertex_attr(g1, name, index = V(graph), value)
```


At this stage we have a simplified version of the friendship graph at time 1 with some basic attributes attached to it such as age, sex, alcohol, smoking etc.

In line with what we have learnt in class we will try and evalute if those rules hold true in the case of this particular graph.The first hypothesis that I try out is checking if this graph is a case of a small world network.


## Small World Graph 

The basic definition of a small worlsd graph are the following - 

1. Characteristic path length - The number of edges in the shortest path averaged over all the paths.
2. Clustering Coefficient - probability that two friends of a ransomly chosen person are friends.

If the graph is a small world (as i expect) what I should observe according to theory is that 
 1. Actual path length ~ Random path length
 2. Actual clustering coeff >> Random clustering coeff
 
 First some exploratory analysis-
 The graph disply structure- sorted by gender
 
 
## Node level statistics

### Computing the indegree and out degree of each node
```{r}

deg_in <- as.data.frame(  degree(g1, mode="in") )
colnames(deg_in) = "deg_in"
deg_out <- as.data.frame(degree(g1, mode="out") )
colnames(deg_out) = "deg_out"
library(ggplot2)

ggplot(data = deg_in, aes(deg_in))+geom_bar()
ggplot(data = deg_out, aes(deg_out))+geom_bar()
```
  There is a strong hint here that there is a large variability in the indegee and out degree of students. A few students are very popular while a majority of studnts have about a median of 3 friends. When we observe outdegree we notice much lesser variability which caps of at 6. it could be because the experimenters might have asked the studnets to name only upto 6 students as their friends. 
 
 
### Computing the shortest path length of each node

We could calculate seperate shortest path lengths using only in or out metrics. I have chosen to calcualte the average of the two 
```{r}
# sp_in <- as.data.frame(  shortest.paths(g1, mode="in") )
# colnames(sp_in) = "sp_in"
# 
# ggplot(data = sp_in, aes(sp_in))+geom_bar()


sp_in <- as.data.frame(  shortest.paths(g1) )
colnames(sp_in) = "sp_in"
ggplot(data = sp_in, aes(sp_in))+geom_bar()

sp_in = do.call(data.frame,lapply(sp_in, function(x) replace(x, is.infinite(x),0)))

avg_path_len = sum(sp_in)/(160*160)
clust_coeff = transitivity(g1)
```
 
## Generating random graph

#R code for generating random graphs:
#requires packages ergm, intergraph

```{r} 
library(ergm)
library(intergraph)



#compute triad count and triad proportion for a given weighting vector
calc_clustering <- function(my.graph){
  n.nodes <- vcount(my.graph)
  n.edges <- ecount(my.graph)
  #set probability of edge formation in random graph to proportion of possible edges present in original
  p.edge <- n.edges/(n.nodes*(n.nodes +1)/2)
  r.graph <- as.network(n.nodes, density = p.edge)
  r.igraph <- asIgraph(r.graph)
  # tc.graph <- triad.census(r.igraph)
  # clust <- sum(tc.graph)
  # clust.norm <- clust/sum(tc.graph)
  clust = transitivity(r.igraph, isolates = "zero" )
  #path = shortest.paths(r.igraph)
  return(c(clust))
}

#build 100 random graphs and compute their clustering and hierarchy measurements to create an empirical null distribution
cluster <- function(this.graph){
  clust <- matrix(rep(0,100), nrow=1) 
  #path <- matrix(rep(0,200),nrow=2)
  for(i in c(1:100)){
    clust[,i] <- calc_clustering(this.graph)
    #path[,i] <- mask.stat(this.graph)
  }
  
  cluster = as.data.frame(t(clust))
}

```

```{r}

######similar calculations for shortest path length

calc_path <- function(my.graph){
  n.nodes <- vcount(my.graph)
  n.edges <- ecount(my.graph)
  #set probability of edge formation in random graph to proportion of possible edges present in original
  p.edge <- n.edges/(n.nodes*(n.nodes +1)/2)
  r.graph <- as.network(n.nodes, density = p.edge)
  r.igraph <- asIgraph(r.graph)
  # tc.graph <- triad.census(r.igraph)
  # clust <- sum(tc.graph)
  # clust.norm <- clust/sum(tc.graph)
  #clust = transitivity(r.igraph, isolates = "zero" )
  path = mean(shortest.paths(r.igraph))
  return(c(path))
}

#build 100 random graphs and compute their clustering and hierarchy measurements to create an empirical null distribution
path <- function(this.graph){
  path <- matrix(rep(0,100), nrow=1) 
  #path <- matrix(rep(0,200),nrow=2)
  for(i in c(1:100)){
    path[,i] <- calc_path(this.graph)
    #path[,i] <- mask.stat(this.graph)
  }
 
  path = as.data.frame(t(path))
}

```

```{r}
#compute empirical distributions for each network
rand_clus_dist <- cluster(g1)
rand_path_dist <- path(g1)


```



Vizualizing the results-

## Avg path length is comparable to random path length

```{r}


ggplot(rand_path_dist, aes(x=V1)) + 
    geom_histogram(aes(y=..density..),binwidth = 0.01,     
                   colour="black", fill="white") +
    geom_density(alpha=.4, fill="blue")

ggplot(rand_clus_dist, aes(x=V1)) + 
    geom_histogram(aes(y=..density..),binwidth = 0.003,     
                   colour="black", fill="white") +
    geom_density(alpha=.4, fill = "red")

```

With only 100 trails we dont have a perfectly normal distribution but it almost seems to work. (with more and more trails the distribution of these values will approch a normal distribution)

Comparing results we have 

| Data                   | Clustering Coefficient   | Avg Path Length |
|------------------------|--------------------------|-----------------|
| School Social Network  | 0.3579                   | 3.7117          |
| Simulated random Graph | 0.0887                   | 2.1543          |





























 
 
 
 How to calculate random path length and random clustering coefficient ? Generating a ER model to draw comparisons . In this case I will use the same degree distributions??? 
 
 





































##Future work and Improvement

1. In line with the paper (Configuring Random Graph Models with Fixed Degree Sequences) choose a better null model which might fit the scenario better. I chose ER graphs as my null model due to my familirity with dealing with them. May not be the best choice theortically.

2. Will the results change if I use a null model with same degree distribution? How robust are the results to different null models. 


