---
title: "Prelim_SNA"
author: "Manu"
date: "February 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading and vizulaizing the data sources:

The objective of the study is to move beyond linear, logistic regression and leverage the network structure to make predictions about peoples behaviour and choices. At baseline we can compare predictions from various parametric models with my proposed non- parametric modelling methodology. 


Simplifications made connections that did not exist (because the students moved or came in from different schools have been counted as null relationships i.e 0). The graph will be directed and weighted. 

The values are recoded as :

0- No friendship
1- Just a friend 
2- Best friend

For now we will use a non weighted directed graph . Just the presence of a friendship will be denoted as 1.

G1 G2 and G3 are the three friendship graphs at three points in time. 
```{r}
library(igraph)

load("C:/Users/ms52/AppData/Local/Temp/Temp1_Glasgow_data.zip/Glasgow-friendship.RData")

# try and viz the friendship networks - 

friendship.1 = ifelse( friendship.1 == 10, 0, friendship.1)
friendship.1 = ifelse( friendship.1 == 2, 1, friendship.1)
g1 <- graph_from_adjacency_matrix(friendship.1)


friendship.2 = ifelse( friendship.2 == 10, 0, friendship.2)
friendship.2 = ifelse( friendship.2 == 2, 1, friendship.2)
g2 <- graph_from_adjacency_matrix(friendship.2)


friendship.3 = ifelse( friendship.3 == 10, 0, friendship.3)
friendship.3 = ifelse( friendship.3 == 2, 1, friendship.3)
g3 <- graph_from_adjacency_matrix(friendship.3)

```
Attach all the various attributes to the friendship graph-

```{r}

plot(g1, edge.arrow.size=.5, vertex.color="gold", vertex.size=15, 

     vertex.frame.color="gray", vertex.label.color="black", 

     vertex.label.cex=0.5, vertex.label.dist=0, edge.curved=0.2) 

 g1 <- graph_from_adjacency_matrix(friendship.1)
 summary(g1)
alcohol = as.data.frame(alcohol)
alcohol$id = rownames(alcohol)
tobacco = as.data.frame(tobacco)
tobacco$id = rownames(tobacco)
cannabis = as.data.frame(cannabis)
cannabis$id = rownames(cannabis)
age = as.data.frame(age)
age$id = rownames(age)
sex.F = as.data.frame(sex.F)
sex.F$id = rownames(sex.F)
dist.school = as.data.frame(dist.school)
dist.school$id = rownames(dist.school)
distance.1 = as.data.frame(distance.1)
distance.1$id = rownames(distance.1)
leisure1 = as.data.frame(leisure1)
leisure1$id = rownames(leisure1)
music1 = as.data.frame(music1)
music1$id = rownames(music1)
familysmoking = as.data.frame(familysmoking)
familysmoking$id = rownames(familysmoking)
romantic = as.data.frame(romantic)
romantic$id = rownames(romantic)
money = as.data.frame(money)
money$id = rownames(money)

other = Reduce(function(x, y) merge(x, y, all=TRUE), list(leisure1,music1 , familysmoking))


# alcohol.1 = alcohol[,c("t1", "id")]
# colnames(alcohol.1) = c("alcohol", "id")
# 
# vertex_attr(g1, index=alcohol$id) <- alcohol.1
# V(net)$Sex=as.character(a$Sex[match(V(net)$name,a$Bird.ID)])
V(g1)$alcohol=as.numeric(alcohol$t1[match(V(g1)$name,alcohol$id)])
V(g1)$tobacco=as.numeric(tobacco$t1[match(V(g1)$name,tobacco$id)])
V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
V(g1)$age=as.numeric(age$t1[match(V(g1)$name,age$id)])
V(g1)$sex=as.numeric(sex.F$t1[match(V(g1)$name,sex.F$id)])
V(g1)$dist.school=as.numeric(dist.school$t1[match(V(g1)$name,dist.school$id)])
V(g1)$distance.1=as.numeric(distance.1$t1[match(V(g1)$name,distance.1$id)])
V(g1)$smoking.at.home=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])
V(g1)$cannabis=as.numeric(cannabis$t1[match(V(g1)$name,cannabis$id)])


g1 = set_vertex_attr(g1, name, index = V(graph), value)
```











